---
title: "ProblemSet1"
format: html
editor: visual
---

## Problem 1 - Wine Data

a\. Import the data

We first import data into the environment. Click the "import Dataset" button and select the file "wine.data". Then we get the data.frame below.

```{r}
wine <- read.csv("wine.data", header = FALSE)
#wine
```

We can also check the format of it.

```{r}
class(wine)
```

Next, we use the information in the "wine.names" file to give column names. We name the first column "class", and the names of following columns are given in the "wine.names" file.

```{r}
colnames(wine)[1] <- "class"
colnames(wine)[2:14] <- c("alcohol", "malic acid", "ash",
                          "Alcalinity of ash", "Magnesium",
                          "Total phenols", "Flavanoids",
                          "Nonflavanoid phenols",
                          "Proanthocyanins", "Color intensity",
                          "Hue", "OD280/OD315 of diluted wines",
                          "Proline")
```

b\. The number of wines within each class

The numbers of wines within the three classes are respectively as follows.

```{r}
# the number of wines within class 1
length(which(wine$class == 1))
# the number of wines within class 2
length(which(wine$class == 2))
# the number of wines within class 3
length(which(wine$class == 3))
```

c\. Use the data to answer the questions

1.  The wine with the highest alcohol content belongs to which class?

    We first find the index of the wine, then find which class it belongs to.

    ```{r}
    index1 <- which.max(wine$alcohol)
    wine$class[index1]
    ```

    Clearly, it belongs to class 1.

2.  The wine with the lowest alcohol content belongs to which class?

    Similarly, we first find the index, then the class it belongs to.

    ```{r}
    index2 <- which.min(wine$alcohol)
    wine$class[index2]
    ```

    It belongs to class 2.

3.  How many of these wines have higher levels of magnesium than 114 mg/l?

    ```{r}
    length(which(wine$Magnesium > 114))
    ```

    There are 26 wines with level of magnesium higher than 114 mg/l.

4.  Within each class, how many wines have higher levels of magnesium than average German beer?

    The number of wines with higher levels of magnesium than average German beer within each class is as follows.

    ```{r}
    # within class 1
    length(which(wine$class == 1 & wine$Magnesium > 114))
    # within class 2
    length(which(wine$class == 2 & wine$Magnesium > 114))
    # within class 3
    length(which(wine$class == 3 & wine$Magnesium > 114))
    ```

d\. Create a table identifying the average value of each variable

We first calculate the overall mean value with respect to each variable.

```{r}
average_overall <- colMeans(wine[2:14])
```

Then we calculate the mean value of each class with respect to each variable.

```{r}
average_class1 <- colMeans(wine[1:59, 2:14])
average_class2 <- colMeans(wine[60:130, 2:14])
average_class3 <- colMeans(wine[131:178 , 2:14])
```

And finally, we combine these mean values together to form a table.

```{r}
average_table <- cbind.data.frame(average_overall, average_class1, average_class2, average_class3)
average_table <- t(average_table)
```
